---
title: "SNAG: T1 and T2 subfields"
author: "Shuying Yu"
date: "1/10/2022"
output: 
  html_document: 
    toc: yes
    toc_float: true
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center")

#Attach libraries
library(tidyverse)
library(here)
library(janitor)
library(ggpubr)
library(patchwork)
```


# Overview

This report provides exploratory and preliminary correlation analyses between sex hormones and T1 and T2 hippocampal subfield data from the Healthy Aging Study sample (up to participant 383). No navigation behavioral data included (see prior analysis or PowerPoint).

I cleaned some of the data prior to analyses. I had to:

1) Filter participants that have either T1 or T2 subfield data from scan
2) Select sex hormone measures from when scan was conducted
3) Correct T1 and T2 subfield volumes using TIV from VBM
4) Average left and right volumes to get an average measure for each subfield and brain region


```{r data_wrangling}
########## Data 

#Read in the data
df <- read_csv(here("data", "snag_t1_t2hipp_hormones.csv"))



#Filter for having T1 and scan hormone data
df_t1 <- df %>% 
  
  #Filter
  filter(t1_status == "yes") %>% 
  
  #Select certain columns
  select(subject_id, sex, age_scan_years,
         repo_status_scan, t1_vol_left_hipp:t1_csf_vbm,
         estradiol_scan_pg_ml:testosterone_scan_ng_dl)


  
df_t2hipp <- df %>% 
  
  #Filter
  filter(t2hipp_status == "yes") %>% 
  
  #Select certain columns
  select(subject_id, sex, age_scan_years,
         repo_status_scan, t1_tiv_vbm,
         t2hipp_tiv_thick_cat12:t2hipp_thick_right_phg,
         estradiol_scan_pg_ml:testosterone_scan_ng_dl)



########## Correct using TIV, then average left and right

##### T1
df_t1_corrected <- df_t1 %>% 

  #Correct using TIV from VBM
  #(Total) hippocampus, MTL, parahippocampal gyrus
  #Include total GMV from hippocampus 
  mutate(t1_vol_left_hipp = t1_vol_left_hipp/t1_tiv_vbm,
         t1_vol_right_hipp = t1_vol_left_hipp/t1_tiv_vbm,
         t1_vol_left_mtl = t1_vol_left_mtl/t1_tiv_vbm,
         t1_vol_right_mtl = t1_vol_right_mtl/t1_tiv_vbm,
         t1_vol_left_parahipp = t1_vol_left_parahipp/t1_tiv_vbm,
         t1_vol_right_parahipp = t1_vol_right_parahipp/t1_tiv_vbm,
         t1_vol_gmv = t1_gmv_vbm/t1_tiv_vbm) %>% 
  
  #Average left and right regions
  mutate(t1_vol_hipp = (t1_vol_left_hipp + t1_vol_right_hipp)/2,
         t1_vol_mtl = (t1_vol_left_mtl + t1_vol_right_mtl)/2,
         t1_vol_parahipp = (t1_vol_left_parahipp + t1_vol_right_parahipp)/2)



##### T2hipp

#Create function for apply to variables
dividebyTIV <- function(x, na.rm = FALSE) (x/df_t2hipp$t1_tiv_vbm)

#Apply function to variables
df_t2hipp_corrected <- df_t2hipp %>% 
  
  #Correct using TIV from T1 VBM
  #CA1, CA23, DG, ERC,PHC, PRC, SUB, PHG
  mutate_at(vars(t2hipp_vol_left_ca1:t2hipp_vol_right_phg),
            dividebyTIV) %>% 
  
  #Average left and right regions
  mutate(t2hipp_vol_ca1 = (t2hipp_vol_left_ca1 + t2hipp_vol_right_ca1)/2,
         t2hipp_vol_ca23 = (t2hipp_vol_left_ca23 + t2hipp_vol_right_ca23)/2,
         t2hipp_vol_dg = (t2hipp_vol_left_dg + t2hipp_vol_right_dg)/2,
         t2hipp_vol_erc = (t2hipp_vol_left_erc + t2hipp_vol_right_erc)/2,
         t2hipp_vol_phc = (t2hipp_vol_left_phc + t2hipp_vol_right_phc)/2,
         t2hipp_vol_sub = (t2hipp_vol_left_sub + t2hipp_vol_right_sub)/2,
         t2hipp_vol_phg = (t2hipp_vol_left_phg + t2hipp_vol_right_phg)/2)
```



# T1 Volume

## Total Hippocampus

```{r}
#Scatterplot
t1hipp1 <- ggscatter(df_t1_corrected, x = "estradiol_scan_pg_ml", y = "t1_vol_hipp",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Total Hippocampus")

#Scatterplot
t1hipp2 <- ggscatter(df_t1_corrected, x = "progesterone_scan_ng_ml", y = "t1_vol_hipp",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Total Hippocampus")


#Scatterplot
t1hipp3 <- ggscatter(df_t1_corrected, x = "fsh_scan_miu_ml", y = "t1_vol_hipp",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Total Hippocampus")


#Scatterplot, facet by sex
t1hipp4 <- ggscatter(df_t1_corrected, x = "shbg_scan_nmol_l", y = "t1_vol_hipp",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Total Hippocampus")

#Scatterplot, facet by sex
t1hipp5 <- ggscatter(df_t1_corrected, x = "testosterone_scan_ng_dl", y = "t1_vol_hipp", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Total Hippocampus")


#Scatterplot, facet by sex
t1hipp6 <- ggscatter(df_t1_corrected, x = "dheas_scan_ug_dl", y = "t1_vol_hipp", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Total Hippocampus")


#Present as patchwork
#(t1hipp1 + t1hipp2 + t1hipp3)/(t1hipp4 + t1hipp5 + t1hipp6)
(t1hipp1 + t1hipp2 + t1hipp3)
t1hipp4
t1hipp5
t1hipp6
```
Total hippocampal volume has a merginal negative association with FSH

## Total GMV

```{r}
#Scatterplot
t1gmv1 <- ggscatter(df_t1_corrected, x = "estradiol_scan_pg_ml", y = "t1_vol_gmv",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Total Grey Matter")

#Scatterplot
t1gmv2 <- ggscatter(df_t1_corrected, x = "progesterone_scan_ng_ml", y = "t1_vol_gmv",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Total Grey Matter")


#Scatterplot
t1gmv3 <- ggscatter(df_t1_corrected, x = "fsh_scan_miu_ml", y = "t1_vol_gmv",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Total Grey Matter")


#Scatterplot, facet by sex
t1gmv4 <- ggscatter(df_t1_corrected, x = "shbg_scan_nmol_l", y = "t1_vol_gmv",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Total Grey Matter")

#Scatterplot, facet by sex
t1gmv5 <- ggscatter(df_t1_corrected, x = "testosterone_scan_ng_dl", y = "t1_vol_gmv", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Total Grey Matter")


#Scatterplot, facet by sex
t1gmv6 <- ggscatter(df_t1_corrected, x = "dheas_scan_ug_dl", y = "t1_vol_gmv", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Total Grey Matter")


#Present as patchwork
(t1gmv1 + t1gmv2 + t1gmv3)
t1gmv4
t1gmv5
t1gmv6
```


## MTL

MTL volume was calculated by adding hippocampal volume (total hippocampus) and parahippocampus volume.

```{r}
#Scatterplot
t1mtl1 <- ggscatter(df_t1_corrected, x = "estradiol_scan_pg_ml", y = "t1_vol_mtl",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Medial Temporal Lobe")

#Scatterplot
t1mtl2 <- ggscatter(df_t1_corrected, x = "progesterone_scan_ng_ml", y = "t1_vol_mtl",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Medial Temporal Lobe")


#Scatterplot
t1mtl3 <- ggscatter(df_t1_corrected, x = "fsh_scan_miu_ml", y = "t1_vol_mtl",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Medial Temporal Lobe")


#Scatterplot, facet by sex
t1mtl4 <- ggscatter(df_t1_corrected, x = "shbg_scan_nmol_l", y = "t1_vol_mtl",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Medial Temporal Lobe")

#Scatterplot, facet by sex
t1mtl5 <- ggscatter(df_t1_corrected, x = "testosterone_scan_ng_dl", y = "t1_vol_mtl", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Medial Temporal Lobe")


#Scatterplot, facet by sex
t1mtl6 <- ggscatter(df_t1_corrected, x = "dheas_scan_ug_dl", y = "t1_vol_mtl", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Medial Temporal Lobe")


#Present as patchwork
(t1mtl1 + t1mtl2 + t1mtl3)
t1mtl4
t1mtl5
t1mtl6
```
MTL volume as a somewhat negative association with FSH



## Parahippocampus

```{r}
#Scatterplot
t1parahipp1 <- ggscatter(df_t1_corrected, x = "estradiol_scan_pg_ml", y = "t1_vol_parahipp",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Parahippocampus")

#Scatterplot
t1parahipp2 <- ggscatter(df_t1_corrected, x = "progesterone_scan_ng_ml", y = "t1_vol_parahipp",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Parahippocampus")


#Scatterplot
t1parahipp3 <- ggscatter(df_t1_corrected, x = "fsh_scan_miu_ml", y = "t1_vol_parahipp",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Parahippocampus")


#Scatterplot, facet by sex
t1parahipp4 <- ggscatter(df_t1_corrected, x = "shbg_scan_nmol_l", y = "t1_vol_parahipp",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Parahippocampus")

#Scatterplot, facet by sex
t1parahipp5 <- ggscatter(df_t1_corrected, x = "testosterone_scan_ng_dl", y = "t1_vol_parahipp", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Parahippocampus")


#Scatterplot, facet by sex
t1parahipp6 <- ggscatter(df_t1_corrected, x = "dheas_scan_ug_dl", y = "t1_vol_parahipp", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Parahippocampus")


#Present as patchwork
(t1parahipp1 + t1parahipp2 + t1parahipp3)
t1parahipp4
t1parahipp5
t1parahipp6
```


# T2 Subfield Volume

## CA1

```{r}
#Scatterplot
t2ca1_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_ca1",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "CA1")

#Scatterplot
t2ca1_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_ca1",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "CA1")


#Scatterplot
t2ca1_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_ca1",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "CA1")


#Scatterplot, facet by sex
t2ca1_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_ca1",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "CA1")

#Scatterplot, facet by sex
t2ca1_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_ca1", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "CA1")


#Scatterplot, facet by sex
t2ca1_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_ca1", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "CA1")


#Present as patchwork
(t2ca1_1 + t2ca1_2 + t2ca1_3)
t2ca1_4
t2ca1_5
t2ca1_6
```

## CA2/3

```{r}
#Scatterplot
t2ca23_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_ca23",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "CA23")

#Scatterplot
t2ca23_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_ca23",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "CA23")


#Scatterplot
t2ca23_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_ca23",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "CA23")


#Scatterplot, facet by sex
t2ca23_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_ca23",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "CA23")

#Scatterplot, facet by sex
t2ca23_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_ca23", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "CA23")


#Scatterplot, facet by sex
t2ca23_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_ca23", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "CA23")


#Present as patchwork
(t2ca23_1 + t2ca23_2 + t2ca23_3)
t2ca23_4
t2ca23_5
t2ca23_6
```



## DG

```{r}
#Scatterplot
t2dg_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_dg",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Dendate Gyrus")

#Scatterplot
t2dg_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_dg",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Dendate Gyrus")


#Scatterplot
t2dg_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_dg",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Dendate Gyrus")


#Scatterplot, facet by sex
t2dg_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_dg",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Dendate Gyrus")

#Scatterplot, facet by sex
t2dg_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_dg", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Dendate Gyrus")


#Scatterplot, facet by sex
t2dg_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_dg", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Dendate Gyrus")


#Present as patchwork
(t2dg_1 + t2dg_2 + t2dg_3)
t2dg_4
t2dg_5
t2dg_6
```


## ERC

```{r}
#Scatterplot
t2erc_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_erc",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Entorhinal Cortex")

#Scatterplot
t2erc_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_erc",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Entorhinal Cortex")


#Scatterplot
t2erc_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_erc",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Entorhinal Cortex")


#Scatterplot, facet by sex
t2erc_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_erc",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Entorhinal Cortex")

#Scatterplot, facet by sex
t2erc_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_erc", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Entorhinal Cortex")


#Scatterplot, facet by sex
t2erc_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_erc", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Entorhinal Cortex")


#Present as patchwork
(t2erc_1 + t2erc_2 + t2erc_3)
t2erc_4
t2erc_5
t2erc_6
```


## PHC

```{r}
#Scatterplot
t2phc_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_phc",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Perirhinal Cortex")

#Scatterplot
t2phc_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_phc",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Perirhinal Cortex")


#Scatterplot
t2phc_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_phc",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Perirhinal Cortex")


#Scatterplot, facet by sex
t2phc_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_phc",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Perirhinal Cortex")

#Scatterplot, facet by sex
t2phc_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_phc", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Perirhinal Cortex")


#Scatterplot, facet by sex
t2phc_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_phc", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Perirhinal Cortex")


#Present as patchwork
(t2phc_1 + t2phc_2 + t2phc_3)
t2phc_4
t2phc_5
t2phc_6
```

## SUB

```{r}
#Scatterplot
t2sub_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_sub",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Subiculum")

#Scatterplot
t2sub_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_sub",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Subiculum")


#Scatterplot
t2sub_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_sub",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Subiculum")


#Scatterplot, facet by sex
t2sub_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_sub",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Subiculum")

#Scatterplot, facet by sex
t2sub_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_sub", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Subiculum")


#Scatterplot, facet by sex
t2sub_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_sub", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Subiculum")


#Present as patchwork
(t2sub_1 + t2sub_2 + t2sub_3)
t2sub_4
t2sub_5
t2sub_6
```


## PHG

```{r}
#Scatterplot
t2phg_1 <- ggscatter(df_t2hipp_corrected, x = "estradiol_scan_pg_ml", y = "t2hipp_vol_phg",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Parahippocampal Gyrus")

#Scatterplot
t2phg_2 <- ggscatter(df_t2hipp_corrected, x = "progesterone_scan_ng_ml", y = "t2hipp_vol_phg",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Parahippocampal Gyrus")


#Scatterplot
t2phg_3 <- ggscatter(df_t2hipp_corrected, x = "fsh_scan_miu_ml", y = "t2hipp_vol_phg",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Parahippocampal Gyrus")


#Scatterplot, facet by sex
t2phg_4 <- ggscatter(df_t2hipp_corrected, x = "shbg_scan_nmol_l", y = "t2hipp_vol_phg",
          facet.by = "sex",
          add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "SHBG (nmol/L)", ylab = "Parahippocampal Gyrus")

#Scatterplot, facet by sex
t2phg_5 <- ggscatter(df_t2hipp_corrected, x = "testosterone_scan_ng_dl", y = "t2hipp_vol_phg", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Testosterone (ng/dL)", ylab = "Parahippocampal Gyrus")


#Scatterplot, facet by sex
t2phg_6 <- ggscatter(df_t2hipp_corrected, x = "dheas_scan_ug_dl", y = "t2hipp_vol_phg", facet.by = "sex",
          add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE,
          cor.method = "pearson",
          xlab = "DHEAS (ug/dL)", ylab = "Parahippocampal Gyrus")


#Present as patchwork
(t2phg_1 + t2phg_2 + t2phg_3)
t2phg_4
t2phg_5
t2phg_6
```


# Summary


Initial findings include:

- For T1 brain data, total hippocampal and MTL volumes are marginally and negatively associated with FSH
- For CA1 and CA2/3, they are marginally and negatively associated with FSH. CA2/3 is also marginally and negatively associated with testosterone for women
- For ERC, it is marginally and positively associated with testosterone in men


Future directions:

- Finish data collection and re-evaluate analysis with larger and complete sample




