---
title: "Study 2, part 2: T2 Subfields and Hormones Correlations"
author: "Shuying Yu and Caitlin Taylor"
date: "08/24/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    theme: flatly
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE,
                      warning = FALSE, message = FALSE, fig.align = "center")

#Attach libraries
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(ggpubr)
library(patchwork)
library(readxl)
library(knitr)
library(broom)
library(purrr)
```



# Overview

This report provides exploratory and preliminary correlation analyses between sex hormones and T2 *unretouched*, raw hippocampal subfield data from the Healthy Aging Study sample that have completed some portion of the spatial navigation study. 

**Cleaning data steps included (T2 subfields):**

1) Filter participants that have T2 subfield data from scan and completed spatial navigation behavior

2) Select sex hormone measures from when scan was conducted

3) Correct left and right volumes with TIV from T1 data

4) Average left and right volumes to get an average measure for each subfield region


**Calculating Total Hippocampal Volume:**

1) After left and right T2 subfields were corrected for TIV, sum up left and right total hipp (CA1, CA23, DG, SUB)

2) Average left and right volumes to get an average measure 


**Calculating Total MTL:**

1) After left and right T2 subfields were corrected for TIV, sum up left and right MTL (CA1, CA23, DG, SUB, ERC, PHC, PRC)

2) Average left and right volumes to get an average measure 


```{r data_wrangling}
########## Data 

#Read in the data
data <- read_csv(here("data", "HAS_SNAG_mastersheet_tidy.csv"))

#Filter for midlife, has spatial, has t2hipp data
data <- data %>% 
  filter(age_group == "Midlife", 
         t2hipp_status == "Yes",
         completed_spatial == "Yes") %>% 
  dplyr::select(subject_id, sex, 
         age_scan_years:testosterone_spatial_ng_dl,
         t2hipp_vol_left_ca1:t2hipp_vol_right_sub,
         t1_vbm_tiv) %>% 
  mutate(repo_status = fct_relevel(repo_status_spatial, 
                                    c("Pre", "Peri", "Post", "Men"))) %>% 
  
  #Further exclude this subject:
  filter(!subject_id %in% c(
    #"309", #bad participant
    #"316", #vertigo/took Wellbutrin for mood and Lipitor for cholesterol 7 days/week of scan
     "329", #on prednisone 7 days/week of scan
    #"352", #current MDD/depression
     "372" #past neuroschistosomiasis at 24 years old
  ))


########## Correct using TIV, then average left and right

##### T2hipp

#Create function for apply to variables
#t2hipp_vol_tiv
dividebyTIV <- function(x, na.rm = FALSE) (x/data$t1_vbm_tiv)

#Correct left and right volumes with TIV
data <- data %>% 
  
  #Correct using TIV from SPM
  #CA1, CA23, DG, ERC,PHC, PRC, SUB, GMV
  mutate_at(vars(t2hipp_vol_left_ca1:t2hipp_vol_right_sub),
            dividebyTIV) %>% 
  
  #Average left and right subfields
  mutate(avg_ca1 = (t2hipp_vol_left_ca1+t2hipp_vol_right_ca1) / 2,
         avg_ca23 = (t2hipp_vol_left_ca23+t2hipp_vol_right_ca23) / 2,
         avg_dg = (t2hipp_vol_left_dg+t2hipp_vol_right_dg) / 2,
         avg_erc = (t2hipp_vol_left_erc+t2hipp_vol_right_erc) / 2,
         avg_phc = (t2hipp_vol_left_phc+t2hipp_vol_right_phc) / 2,
         avg_prc = (t2hipp_vol_left_prc+t2hipp_vol_right_prc) / 2,
         avg_sub = (t2hipp_vol_left_sub+t2hipp_vol_right_sub) / 2) %>% 
  
  #Calculate total left and right hipp
  mutate(total_l_hipp = (t2hipp_vol_left_ca1 + 
                         t2hipp_vol_left_ca23 +
                         t2hipp_vol_left_dg + 
                         t2hipp_vol_left_sub),
         total_r_hipp = (t2hipp_vol_right_ca1 + 
                         t2hipp_vol_right_ca23 + 
                         t2hipp_vol_right_dg + 
                         t2hipp_vol_right_sub),
         avg_total_hipp = (total_l_hipp + total_r_hipp)/2) %>% 
  
  #Calculate left and right MTL
  mutate(lmtl = (t2hipp_vol_left_ca1 + 
                 t2hipp_vol_left_ca23 +
                 t2hipp_vol_left_dg + 
                 t2hipp_vol_left_sub + 
                 t2hipp_vol_left_erc + 
                 t2hipp_vol_left_phc + 
                 t2hipp_vol_left_prc),
         rmtl = (t2hipp_vol_right_ca1 + 
                 t2hipp_vol_right_ca23 +
                 t2hipp_vol_right_dg + 
                 t2hipp_vol_right_sub + 
                 t2hipp_vol_right_erc + 
                 t2hipp_vol_right_phc + 
                 t2hipp_vol_right_prc),
         avg_mtl = (lmtl + rmtl)/2)


#Subset for women by repo status
data_repo <- data %>% 
  filter(sex == "Female") %>% 
  mutate(repo_status = fct_relevel(repo_status, 
                                    c("Pre", "Peri", "Post")))

#Subset for men
data_men <- data %>% 
  filter(sex == "Male")
```


```{r demo-table, include=TRUE, echo=FALSE}
#Table overall
tab_overall <- data %>% 
  group_by(sex) %>% 
  summarize(n_subject = n(),
            age_mean = mean(age_spatial_years),
            age_Sd = sd(age_spatial_years))

#Table LOOP
tab_loop <- data %>% 
  filter(loop_useable == "Yes") %>% 
  group_by(sex) %>% 
  summarize(n_subject = n(),
            age_mean = mean(age_spatial_years),
            age_Sd = sd(age_spatial_years))

#Table MAZE
tab_maze <- data %>% 
  filter(maze_useable == "Yes") %>% 
  group_by(sex) %>% 
  summarize(n_subject = n(),
            age_mean = mean(age_spatial_years),
            age_Sd = sd(age_spatial_years))


#Table DSP
tab_dsp <- data %>% 
  filter(dsp_useable == "Yes") %>% 
  group_by(sex) %>% 
  summarize(n_subject = n(),
            age_mean = mean(age_spatial_years),
            age_Sd = sd(age_spatial_years))


#Join df based on sex
tab1 <- merge(tab_overall, tab_loop, by = "sex")

#Rename columns
tab1 <- tab1 %>% 
  rename(n_subject_overall = n_subject.x,
         age_mean_overall = age_mean.x,
         age_sd_overall = age_Sd.x,
         n_subject_loop = n_subject.y,
         age_mean_loop = age_mean.y,
         age_sd_loop = age_Sd.y)

#Rest of the merge
tab2 <- merge(tab1, tab_maze, by = "sex")
tab_final <- merge(tab2, tab_dsp, by = "sex")


#Rename columns
tab_final <- tab_final %>% 
  rename(n_subject_maze = n_subject.x,
         age_mean_maze = age_mean.x,
         age_sd_maze = age_Sd.x,
         n_subject_dsp = n_subject.y,
         age_mean_dsp = age_mean.y,
         age_sd_dsp = age_Sd.y)

#Kable presentation (ideal if we had all hormones)
tab_final %>% 
  kable(col.names = c("Sex", "n","Mean Age","SD Age",
                      "n","Mean Age","SD Age",
                      "n","Mean Age","SD Age",
                      "n","Mean Age","SD Age"),
        caption = "Sample size and age for women and men grouped by overall completion of spatial navigation behavior with T2hipp scan.",
        digits = 2,
        align = c("l", rep("c", times = 11))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12) %>%
  add_header_above(c(" " = 1, "Overall" = 3, 
                     "LOOP" = 3, "MAZE" = 3, "DSP" = 3))
```

Above table shows the table of sample size and mean and standard deviation of age by sex for overall group who have completed spatial behavior and have a T2hipp scan.

All sample size for T2hipp x hormone correlations for women are listed in the table above. For men, because some participants did not have all hormones evaluated (e.g, only testosterone available), the sample size for the correlations for that hormone and with T1 hipp volume will be different.

For men, the accurate sample sizes would be:

- LOOP: *n* = 10 DHEAS, *n* = 11 for SHBG and T

- MAZE: *n* = 16 DHEAS, *n* = 17 for SHBG and T

- DSP: *n* = 21 DHEAS, *n* = 22 for SHBG and T

```{r}
##### WOMEN
#LOOP
test <- data_repo %>% 
  filter(loop_useable == "Yes") %>% 
  select(subject_id,
         estradiol_spatial_pg_ml,
         progesterone_spatial_ng_ml,
         fsh_spatial_miu_ml,
         shbg_spatial_nmol_l,
         dheas_spatial_ug_dl,
         testosterone_spatial_ng_dl)
#Sum of rows = sample size for each hormone x T1 correlation
colSums(!is.na(test)) #all n=16

#MAZE
test <- data_repo %>% 
  filter(maze_useable == "Yes") %>% 
  select(subject_id,
         estradiol_spatial_pg_ml,
         progesterone_spatial_ng_ml,
         fsh_spatial_miu_ml,
         shbg_spatial_nmol_l,
         dheas_spatial_ug_dl,
         testosterone_spatial_ng_dl)
#Sum of rows = sample size for each hormone x T1 correlation
colSums(!is.na(test)) #all n=50

#DSP
test <- data_repo %>% 
  filter(dsp_useable == "Yes") %>% 
  select(subject_id,
         estradiol_spatial_pg_ml,
         progesterone_spatial_ng_ml,
         fsh_spatial_miu_ml,
         shbg_spatial_nmol_l,
         dheas_spatial_ug_dl,
         testosterone_spatial_ng_dl)
#Sum of rows = sample size for each hormone x T1 correlation
colSums(!is.na(test)) #all n=35


#####MEN
#LOOP
test <- data_men %>% 
  filter(loop_useable == "Yes") %>% 
  select(fsh_spatial_miu_ml,
         shbg_spatial_nmol_l,
         dheas_spatial_ug_dl,
         testosterone_spatial_ng_dl)
#Sum of rows = sample size for each hormone x T1 correlation
colSums(!is.na(test)) #DHEAS n = 10, SHBG & T n = 11

#MAZE
test <- data_men %>% 
  filter(maze_useable == "Yes") %>% 
  select(subject_id,
         shbg_spatial_nmol_l,
         dheas_spatial_ug_dl,
         testosterone_spatial_ng_dl)
#Sum of rows = sample size for each hormone x T1 correlation
colSums(!is.na(test)) #DHEAS n = 16, SHBG & T n = 17

#DSP
test <- data_men %>% 
  filter(dsp_useable == "Yes") %>% 
  select(subject_id,
         shbg_spatial_nmol_l,
         dheas_spatial_ug_dl,
         testosterone_spatial_ng_dl)
#Sum of rows = sample size for each hormone x T1 correlation
colSums(!is.na(test)) #DHEAS n = 21, SHBG & T n = 22
```



# Structural Distributions 

```{r shapiro-tests, echo=FALSE, results='hide'}
#is DV significantly difference from normal?


#TotalHipp
shapiro.test(data$avg_total_hipp) #yes
shapiro.test(data_repo$avg_total_hipp) #yes
shapiro.test(data_men$avg_total_hipp) #no

#MTL
shapiro.test(data$avg_mtl) #yes
shapiro.test(data_repo$avg_mtl) #yes
shapiro.test(data_men$avg_mtl) #no


#CA1
shapiro.test(data$avg_ca1) #no
shapiro.test(data_repo$avg_ca1) #no
shapiro.test(data_men$avg_ca1) #no

#CA23
shapiro.test(data$avg_ca23) #no
shapiro.test(data_repo$avg_ca23) #no
shapiro.test(data_men$avg_ca23) #yes

#DG
shapiro.test(data$avg_dg) #no
shapiro.test(data_repo$avg_dg) #no
shapiro.test(data_men$avg_dg) #no

#ERC
shapiro.test(data$avg_erc) #no
shapiro.test(data_repo$avg_erc) #no
shapiro.test(data_men$avg_erc) #no

#PHC
shapiro.test(data$avg_phc) #no
shapiro.test(data_repo$avg_phc) #no
shapiro.test(data_men$avg_phc) #no

#PRC
shapiro.test(data$avg_prc) #no
shapiro.test(data_repo$avg_prc) #no
shapiro.test(data_men$avg_prc) #no

#SUB
shapiro.test(data$avg_sub) #yes
shapiro.test(data_repo$avg_sub) #yes
shapiro.test(data_men$avg_sub) #no
```

**Structure:** Shapiro-Wilk's test reveal that total hipp and MTL for overall and for women were not normally distributed. Hormone and subfield/volume associations for these subfields will be using Spearman rank correlations.

**Subfields:** Shapiro-Wilk's test reveal that CA23 for men, and SUB overall and SUB for women were not normally distributed. Hormone and subfield/volume associations for these subfields will be using Spearman rank correlations.




# Hormone Distributions 

```{r echo=FALSE, results='hide'}
#RAW

#Estradiol
shapiro.test(data$estradiol_spatial_pg_ml) #yes
shapiro.test(data_repo$estradiol_spatial_pg_ml) #yes

#Progesterone
shapiro.test(data$progesterone_spatial_ng_ml) #yes
shapiro.test(data_repo$progesterone_spatial_ng_ml) #yes

#FSH
shapiro.test(data$fsh_spatial_miu_ml) #yes
shapiro.test(data_repo$fsh_spatial_miu_ml) #yes

#DHEAS
shapiro.test(data$dheas_spatial_ug_dl) #yes
shapiro.test(data_repo$dheas_spatial_ug_dl) #yes
shapiro.test(data_men$dheas_spatial_ug_dl) #no

#SHBG
shapiro.test(data$shbg_spatial_nmol_l) #yes
shapiro.test(data_repo$shbg_spatial_nmol_l) #yes
shapiro.test(data_men$shbg_spatial_nmol_l) #yes

#Testosterone
shapiro.test(data$testosterone_spatial_ng_dl) #yes
shapiro.test(data_repo$testosterone_spatial_ng_dl) #yes
shapiro.test(data_men$testosterone_spatial_ng_dl) #no



#################################################3

#### LOGGED

#Estradiol
shapiro.test(log(data$estradiol_spatial_pg_ml)) #no
shapiro.test(log(data_repo$estradiol_spatial_pg_ml)) #no

#Progesterone
shapiro.test(log(data$progesterone_spatial_ng_ml)) #no
shapiro.test(log(data_repo$progesterone_spatial_ng_ml)) #no

#FSH
shapiro.test(log(data$fsh_spatial_miu_ml)) #yes
shapiro.test(log(data_repo$fsh_spatial_miu_ml)) #yes

#DHEAS
shapiro.test(log(data$dheas_spatial_ug_dl)) #no
shapiro.test(log(data_repo$dheas_spatial_ug_dl)) #no
shapiro.test(log(data_men$dheas_spatial_ug_dl)) #no

#SHBG
shapiro.test(log(data$shbg_spatial_nmol_l)) #no
shapiro.test(log(data_repo$shbg_spatial_nmol_l)) #no
shapiro.test(log(data_men$shbg_spatial_nmol_l)) #no

#Testosterone
shapiro.test(log(data$testosterone_spatial_ng_dl)) #yes
shapiro.test(log(data_repo$testosterone_spatial_ng_dl)) #yes
shapiro.test(log(data_men$testosterone_spatial_ng_dl)) #no
```

Shapiro-Wilk tests reveal that all six hormones violated normality in the overall and women group. Most hormones for women subgroup distributions were still significantly different from normal after log transformation, so Spearman rank correlations will be applied to these hormone and subfield and structure associations. For men, only SHBG violated normality so all subfield and structure correlations with this hormone in men will be spearman rank.


```{r}
## Reproductive Status

### Repo stage: all histograms right-skewed, so apply transformation
h1 <- ggplot(data = data_repo, aes(x = estradiol_spatial_pg_ml)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Estradiol (pg/mL)",
       y = "Count",
       title = "Raw")
h1a <- ggplot(data = data_repo, aes(x = log(estradiol_spatial_pg_ml))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log Estradiol (pg/mL)",
       y = "Count",
       title = "Log Transformed")
h2 <- ggplot(data = data_repo, aes(x = progesterone_spatial_ng_ml)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Progesterone (ng/mL)",
       y = "Count",
       title = "Raw")
h2a <- ggplot(data = data_repo, aes(x = log(progesterone_spatial_ng_ml))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log Progesterone (ng/mL)",
       y = "Count",
       title = "Log Transformed")
h3 <- ggplot(data = data_repo, aes(x = fsh_spatial_miu_ml)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "FSH (mIU/mL)",
       y = "Count",
       title = "Raw")
h3a <- ggplot(data = data_repo, aes(x = log(fsh_spatial_miu_ml))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log FSH (mIU/mL)",
       y = "Count",
       title = "Log Transformed")
h4 <- ggplot(data = data_repo, aes(x = dheas_spatial_ug_dl)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "DHEAS (ug/dL)",
       y = "Count",
       title = "Raw")
h4a <- ggplot(data = data_repo, aes(x = log(dheas_spatial_ug_dl))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log DHEAS (ug/dL)",
       y = "Count",
       title = "Log Transformed")
h5 <- ggplot(data = data_repo, aes(x = shbg_spatial_nmol_l)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "SHBG (nmol/L)",
       y = "Count",
       title = "Raw")
h5a <- ggplot(data = data_repo, aes(x = log(shbg_spatial_nmol_l))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log SHBG (nmol/L)",
       y = "Count",
       title = "Log Transformed")
h6 <- ggplot(data = data_repo, aes(x = testosterone_spatial_ng_dl)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Testosterone (ng/dL)",
       y = "Count",
       title = "Raw")
h6a <- ggplot(data = data_repo, aes(x = log(testosterone_spatial_ng_dl))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log Testosterone (ng/dL)",
       y = "Count",
       title = "Log Transformed")

(h1+h1a)/(h2+h2a)
(h3+h3a)/(h4+h4a)
(h5+h5a)/(h6+h6a)






### Men and women: all right skewed, so transform
h4 <- ggplot(data = data, aes(x = dheas_spatial_ug_dl)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "DHEAS (ug/dL)",
       y = "Count",
       title = "Raw")
h4a <- ggplot(data = data, aes(x = log(dheas_spatial_ug_dl))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log DHEAS (ug/dL)",
       y = "Count",
       title = "Log Transformed")
h5 <- ggplot(data = data, aes(x = shbg_spatial_nmol_l)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "SHBG (nmol/L)",
       y = "Count",
       title = "Raw")
h5a <- ggplot(data = data, aes(x = log(shbg_spatial_nmol_l))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log SHBG (nmol/L)",
       y = "Count",
       title = "Log Transformed")
h6 <- ggplot(data = data, aes(x = testosterone_spatial_ng_dl)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Testosterone (ng/dL)",
       y = "Count",
       title = "Raw")
h6a <- ggplot(data = data, aes(x = log(testosterone_spatial_ng_dl))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log Testosterone (ng/dL)",
       y = "Count",
       title = "Log Transformed")

h4+h4a
h5+h5a
h6+h6a



### Men: DHEAS was normal, but allow all to be log transformed too
data_men <- data %>% 
  filter(sex == "Male")

h4 <- ggplot(data = data_men, aes(x = dheas_spatial_ug_dl)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "DHEAS (ug/dL)",
       y = "Count",
       title = "Raw")
h4a <- ggplot(data = data_men, aes(x = log(dheas_spatial_ug_dl))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log DHEAS (ug/dL)",
       y = "Count",
       title = "Log Transformed")
h5 <- ggplot(data = data_men, aes(x = shbg_spatial_nmol_l)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "SHBG (nmol/L)",
       y = "Count",
       title = "Raw")
h5a <- ggplot(data = data_men, aes(x = log(shbg_spatial_nmol_l))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log SHBG (nmol/L)",
       y = "Count",
       title = "Log Transformed")
h6 <- ggplot(data = data_men, aes(x = testosterone_spatial_ng_dl)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Testosterone (ng/dL)",
       y = "Count",
       title = "Raw")
h6a <- ggplot(data = data_men, aes(x = log(testosterone_spatial_ng_dl))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Log Testosterone (ng/dL)",
       y = "Count",
       title = "Log Transformed")

h4+h4a
h5+h5a
h6+h6a

# 
# For just men, log transformation will also be applied in the following Pearson correlations. Testosterone correlations among men will be using Spearman rank.
```




```{r log-transform}
#Transform hormones so the hormone correlations would run with Pearson
# data_repo <- data_repo %>% 
#   mutate_at(vars(estradiol_spatial_pg_ml:testosterone_spatial_ng_dl), log)
# 
# data <- data %>% 
#   mutate_at(vars(estradiol_spatial_pg_ml:testosterone_spatial_ng_dl), log)
# 
# df_438 <- df_438 %>% 
#   mutate_at(vars(estradiol_spatial_pg_ml:testosterone_spatial_ng_dl), log)
```

# Behavioral Distributions

```{r}
#Navigation behavior subsets
data_loop <- data %>% filter(loop_useable == "Yes")
data_repo_loop <- data_repo %>% filter(loop_useable == "Yes")
data_men_loop <- data_men %>% filter(loop_useable == "Yes")
data_maze <- data %>% filter(maze_useable == "Yes")
data_repo_maze <- data_repo %>% filter(maze_useable == "Yes")
data_men_maze <- data_men %>% filter(maze_useable == "Yes")
data_dsp <- data %>% filter(dsp_useable == "Yes")
data_repo_dsp <- data_repo %>% filter(dsp_useable == "Yes")
data_men_dsp <- data_men %>% filter(dsp_useable == "Yes")

#LOOP: position error
shapiro.test(data_loop$loop_pe_avg_m) #no
shapiro.test(data_repo_loop$loop_pe_avg_m) #no
shapiro.test(data_men_loop$loop_pe_avg_m) #no

#LOOP: degrees traveled
shapiro.test(data_loop$loop_de_avg_degree) #no
shapiro.test(data_repo_loop$loop_de_avg_degree) #no
shapiro.test(data_men_loop$loop_de_avg_degree) #no

#MAZE: accuracy
shapiro.test(data_maze$maze_accuracy_pct) #yes
shapiro.test(data_repo_maze$maze_accuracy_pct) #yes
shapiro.test(data_men_maze$maze_accuracy_pct) #yes

#MAZE: moves
shapiro.test(data_maze$maze_moves) #yes
shapiro.test(data_repo_maze$maze_moves) #yes
shapiro.test(data_men_maze$maze_moves) #no

#DSP: wayfinding success
shapiro.test(data_dsp$dsp_success_pct) #no
shapiro.test(data_repo_dsp$dsp_success_pct) #no
shapiro.test(data_men_dsp$dsp_success_pct) #no

#DSP: solution index
shapiro.test(data_dsp$dsp_si_true_pct) #yes
shapiro.test(data_repo_dsp$dsp_si_true_pct) #yes
shapiro.test(data_men_dsp$dsp_si_true_pct) #no
```

Shapiro-Wilk's test reveal that MAZE accuracy for all and MAZE moves for overall, women groups were not normally distributed. DSP solution index for overall and women were also not normally distributed. Behavior and volume correlations will be Spearman rank


# Total Hippocampal Volume

## Structural Comparison

```{r include=TRUE, echo=FALSE}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_total_hipp,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5, show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Average Total \nHippocampal Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_total_hipp,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Average Total \nHippocampal Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_total_hipp)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average Total \nHippocampal Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```

Box plot shows the women appear to have greater T2 hippocampal GMV than men (some as potential outliers).


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(n_total_hipp = n(),
            mean_total_hipp = mean(avg_total_hipp, na.rm=TRUE),
            std_total_hipp = sd(avg_total_hipp, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean Total Hipp Volume",
                      "SD Total Hipp Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(n_total_hipp = n(),
            mean_total_hipp = mean(avg_total_hipp, na.rm=TRUE),
            std_total_hipp = sd(avg_total_hipp, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean Total Hipp Volume",
                      "SD Total Hipp Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of total hipp are displayed by reproductive status and by sex.


```{r}
#ANOVA of total hipp by repo stage
aov <- aov(avg_total_hipp ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing T2 hipp GMV by reproductive status was not significant.

```{r}
#T-test of total hipp by sex
ttest <- t.test(avg_total_hipp ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing GMV by sex was significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) have larger T2 hippocampal volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


## Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Average Total\n Hippocampal Volume") +
    theme(legend.position = "none") 


#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Average Total\n Hippocampal Volume") +
    facet_grid(repo_status ~ ., scale = "free")+
     theme_bw() +
     theme(legend.position = "none")


#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Average Total\n Hippocampal Volume") +
    theme(legend.position = "none") 

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Average Total\n Hippocampal Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Average Total\n Hippocampal Volume") +
    theme(legend.position = "none") 

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Average Total\n Hippocampal Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
     theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Average Total\n Hippocampal Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Average Total\n Hippocampal Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Average Total\n Hippocampal Volume") +
    theme(legend.position = "none")


#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Average Total\n Hippocampal Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Average Total\n Hippocampal Volume") +
    theme(legend.position = "none")


#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_total_hipp", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Average Total\n Hippocampal Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman. 

## Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_total_hipp",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged Total\n Hippocampal Volume") +
     theme_bw() +
     theme(legend.position = "none")

#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_total_hipp",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged Total\n Hippocampal Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_total_hipp", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged Total\n Hippocampal Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```

All correlations are Pearson except for SHBG (Spearman)

## Behavioral Correlations {.tabset}

### LOOP

### MAZE

### DSP

## {-}


# MTL

## Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_mtl,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Avergage MTL Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_mtl,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Average MTL Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_mtl)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average MTL Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_mtl = mean(avg_mtl, na.rm=TRUE),
            std_mtl = sd(avg_mtl, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean MTL Volume",
                      "SD MTL Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)


data %>%
  group_by(sex) %>%
  summarize(count = n(),
            mean_mtl = mean(avg_mtl, na.rm=TRUE),
            std_mtl = sd(avg_mtl, na.rm=TRUE)) %>%
  kable(col.names = c("Sex", "n", "Mean MTL Volume",
                      "SD MTL Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>%
  kable_styling("striped", full_width = FALSE,
                position = "center",
                font_size = 12)
```

Means and standard deviations of MTL are displayed by reproductive status and by sex.


```{r}
#ANOVA of MTL by repo stage
aov <- aov(avg_mtl ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing MTL by reproductive status was not significant.

```{r}
#T-test of MTL by sex
ttest <- t.test(avg_mtl ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```


T-test results comparing MTL by sex was significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) have larger MTL volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).



## Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Average MTL Volume") +
    theme(legend.position = "none") 


#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Average MTL Volume") +
    facet_grid(repo_status ~ ., scale = "free")+
     theme_bw() +
     theme(legend.position = "none")


#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Average MTL Volume") +
    theme(legend.position = "none") 

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Average MTL Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Average MTL Volume") +
    theme(legend.position = "none") 

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Average MTL Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
     theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Average MTL Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Average MTL Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Average MTL Volume") +
    theme(legend.position = "none")


#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Average MTL Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Average MTL Volume") +
    theme(legend.position = "none")


#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_mtl", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Average MTL Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```


All correlations are Spearman.


## Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_mtl",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged MTL Volume") +
     theme_bw() +
     theme(legend.position = "none")

#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_mtl",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged MTL Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_mtl", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged MTL Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```


All correlations are Pearson except SHBG (Spearman)

# T2 Subfield Volume {.tabset}

## CA1 

### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_ca1,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged CA1 \nSubfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_ca1,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged CA1 \nSubfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_ca1)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average CA1 Subfield Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_ca1 = mean(avg_ca1, na.rm=TRUE),
            std_ca1 = sd(avg_ca1, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean CA1 Volume",
                      "SD CA1 Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_ca1 = mean(avg_ca1, na.rm=TRUE),
            std_ca1 = sd(avg_ca1, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean CA1 Volume",
                      "SD CA1 Volume"),
        caption = "Sample size and means for subjects by sex status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex


```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_ca1 ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing CA1 subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_ca1 ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```


T-test results comparing CA1 subfield by sex was significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) have larger CA1 volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).

### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged CA1 Subfield Volume") +
    theme(legend.position = "none") 


#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged CA1 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free")+
     theme_bw() +
     theme(legend.position = "none")


#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged CA1 Subfield Volume") +
    theme(legend.position = "none") 

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged CA1 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged CA1 Subfield Volume") +
    theme(legend.position = "none") 

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged CA1 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
     theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged CA1 Subfield Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged CA1 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged CA1 Subfield Volume") +
    theme(legend.position = "none")


#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged CA1 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged CA1 Subfield Volume") +
    theme(legend.position = "none")


#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_ca1", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged CA1 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman.


### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_ca1",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged CA1 Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_ca1",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged CA1 Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_ca1", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged CA1 Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```

Correlations are all Pearson except SHBG (Spearman)



## CA2/3


### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_ca23,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged CA23 \nSubfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_ca23,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged CA23 \nSubfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 

#Total
p3 <- ggplot(data = data, aes(x = avg_ca23)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average CA23 Subfield Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```

```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_ca23 = mean(avg_ca23, na.rm=TRUE),
            std_ca23 = sd(avg_ca23, na.rm=TRUE)) %>%
  kable(col.names = c("Reproductive Status", "n", "Mean CA23 Volume",
                      "SD CA23 Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_ca23 = mean(avg_ca23, na.rm=TRUE),
            std_ca23 = sd(avg_ca23, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean CA23 Volume",
                      "SD CA23 Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12) 
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex


```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_ca23 ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing CA23 subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_ca23 ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing CA23 subfield by sex was significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) have larger CA23 volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged CA23 Subfield Volume") +
    theme(legend.position = "none")

#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged CA23 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") + theme_bw() +
    theme(legend.position = "none")

#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged CA23 Subfield Volume") +
    theme(legend.position = "none")

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged CA23 Subfield Volume") + facet_grid(repo_status ~ ., scale = "free") + theme_bw() +
    theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged CA23 Subfield Volume") +
    theme(legend.position = "none")

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged CA23 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") + theme_bw() +
    theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged CA23 Subfield Volume")+
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged CA23 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged CA23 Subfield Volume")+
    theme(legend.position = "none")

#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged CA23 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")



#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged CA23 Subfield Volume") +
    theme(legend.position = "none")

#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_ca23", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged CA23 Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") + theme_bw() +
    theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```


All correlations were Spearman

### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_ca23",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged CA23 Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_ca23", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged CA23 Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_ca23",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged CA23 Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```

Correlations are all Spearman.

## DG


### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_dg,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged DG \nSubfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_dg,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged DG \nSubfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_dg)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average DG Subfield Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_dg = mean(avg_dg, na.rm=TRUE),
            std_dg = sd(avg_dg, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean DG Volume",
                      "SD DG Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_dg = mean(avg_dg, na.rm=TRUE),
            std_dg = sd(avg_dg, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean DG Volume",
                      "SD DG Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex

```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_dg ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing DG subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_dg ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing DG subfield by sex was not significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) do not have larger DG volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged DG Subfield Volume") +
    theme(legend.position = "none")

#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged DG Subfield Volume") +
     facet_grid(repo_status ~ ., scale = "free") +theme_bw() +
    theme(legend.position = "none")

#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged DG Subfield Volume") +
    theme(legend.position = "none")

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged DG Subfield Volume") +
     facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged DG Subfield Volume") +
    theme(legend.position = "none")

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged DG Subfield Volume") +
     facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged DG Subfield Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged DG Subfield Volume") +
     facet_grid(repo_status ~ ., scale = "free") +
   theme_bw() +
    theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged DG Subfield Volume") +
    theme(legend.position = "none")

#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged DG Subfield Volume") +
     facet_grid(repo_status ~ ., scale = "free") +
   theme_bw() +
    theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged DG Subfield Volume") +
    theme(legend.position = "none")

#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_dg", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged DG Subfield Volume") +
     facet_grid(repo_status ~ ., scale = "free") +
   theme_bw() +
    theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman

### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_dg",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged DG Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_dg",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged DG Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_dg",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged DG Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Present as patchwork
c1+c2+c3
```

All correlations are Pearson except SHBG (Spearman)

## ERC

### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_erc,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged ERC \nSubfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_erc,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged ERC \nSubfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_erc)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average ERC Subfield Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_erc = mean(avg_erc, na.rm=TRUE),
            std_erc = sd(avg_erc, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean ERC Volume",
                      "SD ERC Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_erc = mean(avg_erc, na.rm=TRUE),
            std_erc = sd(avg_erc, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean ERC Volume",
                      "SD ERC Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex

```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_erc ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing ERC subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_erc ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing ERC subfield by sex was significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) have larger ERC volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged ERC Subfield Volume") +
    theme(legend.position = "none")

#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged ERC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged ERC Subfield Volume") +
    theme(legend.position = "none")

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged ERC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged ERC Subfield Volume") +
    theme(legend.position = "none")

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged ERC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged ERC Subfield Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged ERC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged ERC Subfield Volume") +
    theme(legend.position = "none")

#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged ERC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged ERC Subfield Volume") +
    theme(legend.position = "none")

#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_erc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged ERC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman.

### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_erc",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged ERC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_erc", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged ERC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_erc",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged ERC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```
All correlations are Pearson except SHBG (Spearman)


## PHC


### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_phc,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged PHC \nSubfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none")

#Sex
p2 <- ggplot(data = data, aes(y = avg_phc,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged PHC \nSubfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 

#Total
p3 <- ggplot(data = data, aes(x = avg_phc)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average PHC Subfield Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_phc = mean(avg_phc, na.rm=TRUE),
            std_phc = sd(avg_phc, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean PHC Volume",
                      "SD PHC Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_phc = mean(avg_phc, na.rm=TRUE),
            std_phc = sd(avg_phc, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean PHC Volume",
                      "SD PHC Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex

```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_phc ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing PHC subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_phc ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing PHC subfield by sex was not significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) do not have larger PHC volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged PHC Subfield Volume") +
    theme(legend.position = "none")

#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged PHC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged PHC Subfield Volume") +
    theme(legend.position = "none")

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged PHC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged PHC Subfield Volume") +
    theme(legend.position = "none")

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged PHC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged PHC Subfield Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged PHC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged PHC Subfield Volume") +
    theme(legend.position = "none")

#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged PHC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged PHC Subfield Volume") +
    theme(legend.position = "none")

#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_phc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged PHC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman

### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_phc", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged PHC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_phc", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged PHC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_phc", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged PHC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Present as patchwork
c1+c2+c3
```

All correlations are Pearson except SHBG (Spearman)

## PRC


### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_prc,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged PRC Subfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none") 

#Sex
p2 <- ggplot(data = data, aes(y = avg_prc,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged PRC Subfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_prc)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average PRC Subfield Volume",
       y = "Count",
       title = "Total")


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_prc = mean(avg_prc, na.rm=TRUE),
            std_prc = sd(avg_prc, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean PRC Volume",
                      "SD PRC Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_prc = mean(avg_prc, na.rm = TRUE),
            std_prc = sd(avg_prc, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean PRC Volume",
                      "SD PRC Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex


```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_prc ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing PRC subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_prc ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing PRC subfield by sex was not significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) do not have larger PRC volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged PRC Subfield Volume") +
    theme(legend.position = "none")

#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged PRC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged PRC Subfield Volume") +
    theme(legend.position = "none")

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged PRC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged PRC Subfield Volume") +
    theme(legend.position = "none")

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged PRC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged PRC Subfield Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged PRC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged PRC Subfield Volume") +
    theme(legend.position = "none")

#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged PRC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged PRC Subfield Volume") +
    theme(legend.position = "none")

#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_prc", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged PRC Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman

### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_prc", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged PRC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_prc", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged PRC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_prc",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged PRC Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```

All correlations are Pearson except SHBG (Spearman)


## SUB

### Structural Comparison

```{r}
#Reproductive status
p1 <- ggplot(data = data_repo, aes(y = avg_sub,
                                 x = repo_status,
                                 fill = repo_status,
                                 color = repo_status)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Reproductive Status",
       y = "Averaged SUB \nSubfield Volume",
       title = "Reproductive Status") + 
  theme(legend.position = "none")

#Sex
p2 <- ggplot(data = data, aes(y = avg_sub,
                                 x = sex,
                                 fill = sex,
                                 color = sex)) + 
  geom_boxplot(alpha = 0.5,show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = FALSE) +
  stat_summary(fun.y=mean, geom="point",
               shape=20, size=3, color="black", fill="black") +
  labs(x = "Sex",
       y = "Averaged SUB \nSubfield Volume",
       title = "Sex") + 
  theme(legend.position = "none") 


#Total
p3 <- ggplot(data = data, aes(x = avg_sub)) + 
  geom_histogram(alpha = 0.5, position = "identity",
                 show.legend = FALSE) +
  labs(x = "Average SUB Subfield Volume",
       y = "Count",
       title = "Total") 


#Combine
(p1+p2)/p3
```


```{r include=TRUE, echo=FALSE}
data_repo %>% 
  group_by(repo_status) %>% 
  summarize(count = n(),
            mean_sub = mean(avg_sub, na.rm=TRUE),
            std_sub = sd(avg_sub, na.rm=TRUE)) %>% 
  kable(col.names = c("Reproductive Status", "n", "Mean SUB Volume",
                      "SD SUB Volume"),
        caption = "Sample size and means for subjects by reproductive status with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)

data %>% 
  group_by(sex) %>% 
  summarize(count = n(),
            mean_sub = mean(avg_sub, na.rm=TRUE),
            std_sub = sd(avg_sub, na.rm=TRUE)) %>% 
  kable(col.names = c("Sex", "n", "Mean SUB Volume",
                      "SD SUB Volume"),
        caption = "Sample size and means for subjects by sex with T2 hipp volume and completed spatial navigation behavior",
        digits = 2,
        align = c("l", rep("c", times = 3))) %>% 
  kable_styling("striped", full_width = FALSE,
                position = "center", 
                font_size = 12)
```

Means and standard deviations of the subfield are displayed by reproductive status and by sex

```{r}
#ANOVA of subfield by repo stage
aov <- aov(avg_sub ~ repo_status, data = data_repo)
summary_aov <- summary(aov)
kable(summary_aov[[1]])
```

ANOVA results comparing SUB subfield by reproductive status was not significant.

```{r}
#T-test of subfield by sex
ttest <- t.test(avg_sub ~ sex, data = data)
ttest
tidy_ttest <- tidy(ttest)
```

T-test results comparing SUB subfield by sex was significant. Women (*M* = `r round(tidy_ttest$estimate1, digits = 3)`) have larger SUB volume than men (*M* = `r round(tidy_ttest$estimate2, digits = 3)`) (*t* = `r round(tidy_ttest$statistic, digits = 3)`, *df* = `r round(tidy_ttest$parameter, digits = 3)`, *p* = `r round(tidy_ttest$p.value, digits = 3)`).


### Hormone Correlation by Reproductive Status

```{r}
#Estradiol
c1 <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged SUB Subfield Volume") +
    theme(legend.position = "none")

#Estradiol facet
c1a <- ggscatter(data_repo, x = "estradiol_spatial_pg_ml", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Estradiol (pg/mL)", ylab = "Averaged SUB Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#Progesterone
c2 <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged SUB Subfield Volume") +
    theme(legend.position = "none")

#Progesterone facet
c2a <- ggscatter(data_repo, x = "progesterone_spatial_ng_ml", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Progesterone (ng/mL)", ylab = "Averaged SUB Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#FSH
c3 <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged SUB Subfield Volume") +
    theme(legend.position = "none")

#FSH facet
c3a <- ggscatter(data_repo, x = "fsh_spatial_miu_ml", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "FSH (mIU/mL)", ylab = "Averaged SUB Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")

#SHBG
c4 <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged SUB Subfield Volume") +
    theme(legend.position = "none")

#SHBG facet
c4a <- ggscatter(data_repo, x = "shbg_spatial_nmol_l", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged SUB Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +theme_bw() +
    theme(legend.position = "none")


#DHEAS
c5 <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged SUB Subfield Volume") +
    theme(legend.position = "none")

#DHEAS facet
c5a <- ggscatter(data_repo, x = "dheas_spatial_ug_dl", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged SUB Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +theme_bw() +
    theme(legend.position = "none")


#Testosterone
c6 <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged SUB Subfield Volume") +
    theme(legend.position = "none")

#Testosterone facet
c6a <- ggscatter(data_repo, x = "testosterone_spatial_ng_dl", y = "avg_sub", color = "repo_status",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged SUB Subfield Volume") +
    facet_grid(repo_status ~ ., scale = "free") +
    theme_bw() +
    theme(legend.position = "none")


#Present as patchwork
(c1 + c1a)
(c2 + c2a)
(c3 + c3a)
(c4 + c4a)
(c5 + c5a)
(c6 + c6a)
```

All correlations are Spearman.


### Hormone Correlation by Men

```{r}
#SHBG
c1 <- ggscatter(data_men, x = "shbg_spatial_nmol_l", y = "avg_sub",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "spearman",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "SHBG (nmol/L)", ylab = "Averaged SUB Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#DHEAS
c2 <- ggscatter(data_men, x = "dheas_spatial_ug_dl", y = "avg_sub", 
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "DHEAS (ug/dL)", ylab = "Averaged SUB Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")

#Testosterone
c3 <- ggscatter(data_men, x = "testosterone_spatial_ng_dl", y = "avg_sub",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Testosterone (ng/dL)", ylab = "Averaged SUB Subfield Volume") +
     theme_bw() +
     theme(legend.position = "none")


#Present as patchwork
c1+c2+c3
```

All correlations are Pearson except SHBG (Spearman)

## {-}



# Summary

## Hormones

**Total hipp Volume and MTL**

- Sample included HAS participants who had a T2 hipp scan and completed spatial navigation behavior.

- ANOVA results comparing total hipp volume and MTL by reproductive stage were not significant. T-test results comparing total hipp volume (*p* < 0.01) and MTL (*p* < 0.05) by sex were significant, with results showing women have larger volumes than men.

- For total hipp volume, it was negatively correlated with estradiol for pres (*p* = 0.033, Spearman) and marginally positively correlated with FSH for peris (*p* = 0.07, Spearman).

- There were no significant correlations between MTL and hormones observed.

**Subfields**

- Sample included HAS participants who had a T2 hipp scan and completed spatial navigation behavior.

- Subregion volumes for women and men were generally normally distributed, except for CA23 in men and SUB in women. ANOVA results comparing subfields by reproductive stage were all not significant. ANOVA results for ERC was marginally significant by reproductive stage (*p* = 0.087). T-test results comparing subfields by sex were all significant (at least *p* < 0.05) except for DG, PHC, and PRC, with results revealing that women larger volumes than men.

- All hormones were skewed for overall and women groups, so hormone and volume correlations were Spearman rank correlations. All correlations for men were pearson except for those involving SHBG. CA23 correlations with hormones for men were all Spearman rank.

- For estradiol, it was significantly negatively correlated with with DG for peris (*p* < 0.01, Spearman) and with SUB for pres (*p* = 0.044, Spearman). It was also marginally negatively associated with CA23 for pres (*p* = 0.07, Spearman).

- For progesterone, it was marginally positively associated with ERC for posts (*p* = 0.053, Spearman).

- For FSH, it was significantly negatively correlated with ERC for posts (*p* < 0.01, Spearman) and with SUB for posts (*p* = 0.031, Spearman). FSH is marginally negatively correlated with CA1 for pres (*p* = 0.07, Spearman) and posts (*p* = 0.07, Spearman) and with PHC for women as a group (*p* = 0.08, Spearman). FSH was also marginally positively correlated with CA23 for peris (*p* = 0.08, Spearman) and with ERC for pres (*p* = 0.07, Spearman). 

- For SHBG, it was significantly negatively correlated with ERC for posts (*p* < 0.01, Spearman). It was also significantly negatively correlated with PRC among women as a group (*p* = 0.038, Spearman) and for pres (*p* = 0.04, Spearman). SHBG was also marginally negatively correlated with SUB for women as a group (*p* = 0.07, Spearman).

- For DHEAS, it was significantly negatively correlated with CA23 in women as a group (*p* = 0.01, Spearman).

- For testosterone, it was marginally negatively correlated with CA1 for men (*p* = 0.065, Pearson).


## Behavior





