---
title: 'Study 2: T2 Subfield, Spatial Behavior, Sex Hormones'
author: "Shuying Yu"
date: "8/5/2022"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      include = FALSE)

#Attach libraries
library(tidyverse)
library(here)
library(janitor)
library(ggpubr)
library(patchwork)
library(kableExtra)
library(foreign)
library(RColorBrewer)
library(plotly)
library(readr)
```


# Introduction

Study 1 has shown that there is a correlation between sex hormones and spatial navigation strategy, such that higher estradiol is positively associated with higher use of shortcuts in the DSP task (higher solution index). Higher FSH, on the other hand, was marginally negatively associated with shortcut use (lower solution index) using Spearman rank correlations.

Here, I determine whether there are associations between T2 high-resolution hippocampal subfield volume and spatial navigation behaviors, and sex hormones and subfield volume. Later, I determine how subfield volume mediates the relationship between sex hormones and behavior for navigation strategy.

Prior exploratory analysis for relationships between sex hormones and subfield volume reveal certain significant relationships (`correlation_greymatter_hormones_corrected.Rmd`).

Here, I will explore whether there are any relationships between spatial navigation behaviors and T2 subfield volumes. Next, I plan to apply causal mediation analysis to investigate whether the effect of sex hormones (IV) on spatial behavior (DV) is mediated by participants' subfield volume (mediator, T2 subfield).


# Data Analysis

I cleaned some of the data prior to analyses. I had to:

1) Filter for midlife participants that have T2 subfield data (subject 333 and over)
2) Select sex hormone measures from when scan was conducted
3) Correct for subfield volumes using TIV, which is a combination of GMV, WMV, CSF. Divide the volumes by TIV variable (`t2hipp_vol_tiv`)
4) Average left and right volumes to get an average measure for each subfield and brain region



```{r}
#Read in the data
data <- read_csv(here("data", "HAS_SNAG_mastersheet_tidy.csv"))

#Filter for midlife, has snag data, has t2hipp data
data <- data %>% 
  filter(age_group == "Midlife", 
         completed_spatial == "Yes",
         t2hipp_status == "Yes") %>% 
  select(subject_id, sex, age_scan_years:testosterone_spatial_ng_dl,
         t2hipp_vol_left_ca1:t2hipp_vol_csf) %>% 
  mutate(repo_status_spatial = fct_relevel(repo_status_spatial, 
                                    c("Pre", "Peri", "Post")))

#Create function for apply to variables
dividebyTIV <- function(x, na.rm = FALSE) (x/data$t2hipp_vol_tiv)

#Correct left and right volumes with TIV
data <- data %>% 
  mutate_at(vars(t2hipp_vol_left_ca1:t2hipp_vol_right_sub),
            dividebyTIV)

#Average left and right volumes
data <- data %>% 
  mutate(t2hipp_vol_ca1_cor = (t2hipp_vol_left_ca1+t2hipp_vol_right_ca1) / 2,
         t2hipp_vol_ca23_cor = (t2hipp_vol_left_ca23+t2hipp_vol_right_ca23) / 2,
         t2hipp_vol_dg_cor = (t2hipp_vol_left_dg+t2hipp_vol_right_dg) / 2,
         t2hipp_vol_erc_cor = (t2hipp_vol_left_erc+t2hipp_vol_right_erc) / 2,
         t2hipp_vol_phc_cor = (t2hipp_vol_left_phc+t2hipp_vol_right_phc) / 2,
         t2hipp_vol_prc_cor = (t2hipp_vol_left_prc+t2hipp_vol_right_prc) / 2,
         t2hipp_vol_sub_cor = (t2hipp_vol_left_sub+t2hipp_vol_right_sub) / 2
         # t2hipp_vol_totalhipp_cor = t2hipp_vol_totalhipp / 2,
         # t2hipp_vol_phg_cor = t2hipp_vol_phg / 2,
         # t2hipp_vol_mtl_cor = t2hipp_vol_mtl / 2
         )

#Separate for SNAG
data_loop <- data %>% 
  filter(loop_useable == "Yes")
data_maze <- data %>% 
  filter(maze_useable == "Yes")
data_dsp <- data %>% 
  filter(dsp_useable == "Yes")

#Separate for women and men
data_women <- data %>% 
  filter(sex == "Female", dsp_useable == "Yes") 
data_men <- data %>% 
  filter(sex == "Male", dsp_useable == "Yes")
```


## Correlation for Behavior and T2 Subfields

From prior analyses, we know that the subfield volumes are generally normally distributed. We can check the distribution of the SNAG dependent variables


```{r}
#LOOP: normally distributed, no need to transform
h1 <- ggplot(data = data_loop, aes(x = loop_pe_avg_m)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Average Position Error (m)",
       y = "Count",
       title = "LOOP Raw")
h2 <- ggplot(data = data_loop, aes(x = loop_de_avg_degree)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Average Degrees Traveled",
       y = "Count",
       title = "LOOP Raw")

#MAZE:right and left skewed, log transform
h3 <- ggplot(data = data_maze, aes(x = maze_accuracy_pct)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Wayfinding Success (%)",
       y = "Count",
       title = "MAZE Raw")
h3a <- ggplot(data = data_maze, aes(x = log(maze_accuracy_pct))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Wayfinding Success (%)",
       y = "Count",
       title = "MAZE Log Transformed")
h4 <- ggplot(data = data_maze, aes(x = maze_moves)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Number of Moves",
       y = "Count",
       title = "MAZE Raw")
h4a <- ggplot(data = data_maze, aes(x = log(maze_moves))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Number of Moves",
       y = "Count",
       title = "MAZE Raw")

#DSP: slight left and right skew so apply transformation
h5 <- ggplot(data = data_dsp, aes(x = dsp_success_pct)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Wayfinding Success (%)",
       y = "Count",
       title = "DSP Raw")
h5a <- ggplot(data = data_dsp, aes(x = log(dsp_success_pct))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Wayfinding Success (%)",
       y = "Count",
       title = "DSP Log Transformed")
h6 <- ggplot(data = data_dsp, aes(x = dsp_si_true_pct)) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Solution Index (%)",
       y = "Count",
       title = "DSP Raw")
h6a <- ggplot(data = data_dsp, aes(x = log(dsp_si_true_pct))) +
  geom_histogram(alpha = 0.5, color = "black", position = "identity") +
  labs(x = "Solution Index (%)",
       y = "Count",
       title = "DSP Log Transformed")


#Present
(h1+h2)
(h3+h3a)/(h4+h4a)
(h5+h5a)/(h6+h6a)
```

LOOP dependent variables are normally distributed, but MAZE and DSP variables have slight skews. I applied log transformation to these variables to allow for exploratory Pearson correlations, but skew remains. For MAZE and DSP, I'll use Spearman rank correlations.



```{r log-transformation}
data_maze_log <- data_maze %>% 
  mutate_at(vars(maze_accuracy_pct,maze_moves), log)

data_dsp_log <- data_dsp %>% 
  mutate_at(vars(dsp_success_pct, dsp_si_true_pct), log)
```


### LOOP

Overall (women and men) and by sex

#### Position Error

```{r}
#CA1
c1 <- ggscatter(data_loop, x = "t2hipp_vol_ca1_cor", y = "loop_pe_avg_m", color = "repo_status_spatial",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Averaged CA1 Subfield Volume", ylab = "Averaged \nPosition Error (m)") +
    theme(legend.position = "top") 

c1a <- ggscatter(data_loop, x = "t2hipp_vol_ca1_cor", y = "loop_pe_avg_m", color = "sex",
          add = "reg.line",
          add.params = list(color = "black", fill = "lightgray"), # Customize reg. line
          cor.coef = TRUE, # Add correlation coefficient. see ?stat_cor
          conf.int = TRUE, 
          cor.method = "pearson",
          cor.coeff.args = list(label.sep = "\n"),
          xlab = "Averaged CA1 Subfield Volume", ylab = "Averaged \nPosition Error (m)") +
    facet_grid(. ~ sex, scale = "free") +
     theme_bw() +
     theme(legend.position = "none")

#CA23

#DG

#ERC

#PHC

#PRC

#SUB



#Patchwork
c1/c1a
```


#### Degrees Traveled


### MAZE


#### Wayfinding Success

#### Number of Moves


### DSP

#### Wayfinding Success

#### Solution Index









## Total Effect (Hormones -> Behavior)


```{r}
#Estradiol

#wayfinding success
fit.totaleffect <- lm(dsp_success_pct ~ estradiol_spatial_pg_ml,
                     data = data_women)
summary(fit.totaleffect)

#solution index
fit.totaleffect <- lm(dsp_si_true_pct ~ estradiol_spatial_pg_ml,
                     data = data_women)
summary(fit.totaleffect)
```

There doesn’t necessarily have to be a significant relationship between the iv and the dv. Just as correlation doesn’t prove causation, no correlation does not disprove causation 

## IV on Mediator

```{r}

```








## Subregion Volume and Spatial Behavior


Assessing any relationships between subregion volume and SNAG behavior, specifically with DSP. First, check normality of the volume data

```{r}
#Subregions
par(mfrow=c(2,2))
hist(data_women$t2hipp_vol_ca1_cor)
hist(data_women$t2hipp_vol_ca23_cor)
hist(data_women$t2hipp_vol_mtl_cor)
hist(data_women$t2hipp_vol_totalhipp_cor)


#DSP 
par(mfrow=c(1,2))
hist(data_women$dsp_success_pct)
hist(data_women$dsp_si_true_pct)
```
Subregion volumes appear to be normally distributed. There is an outlier wirh the MTL corrected subregion volume (max 20997.82 compared to mean of 7986.304).

For DSP, wayfinding success if left skewed and solution index is right skewed. I try to see if I can take the log of these dependent variables to normalize them for Pearson correlation

```{r}
#log transform
par(mfrow=c(1,2))
hist(log(data_women$dsp_success_pct))
hist(log(data_women$dsp_si_true_pct))

#inverse
par(mfrow=c(1,2))
hist((data_women$dsp_success_pct)^-1)
hist((data_women$dsp_si_true_pct)^-1)

#square root
par(mfrow=c(1,2))
hist(sqrt(data_women$dsp_success_pct))
hist(sqrt(data_women$dsp_si_true_pct))
```

Log transformation and square root makes the DSP dependent variable more normally distributed, moreso for solution index than for wayfinding success. Use spearman rank correlation instead.


```{r}
#Correlations: wayfinding success and CA1, CS23, MTL, totalhipp

########## Spearman for wayfinding success

#CA1
cor.test(data_women$t2hipp_vol_ca1_cor,
         data_women$dsp_success_pct,
         conf.level = 0.95,
         method = "spearman")

#CA23
cor.test(data_women$t2hipp_vol_ca23_cor,
         data_women$dsp_success_pct,
         conf.level = 0.95,
         method = "spearman")

#MTL
cor.test(data_women$t2hipp_vol_mtl_cor,
         data_women$dsp_success_pct,
         conf.level = 0.95,
         method = "spearman")

#Total hipp
cor.test(data_women$t2hipp_vol_totalhipp_cor,
         data_women$dsp_success_pct,
         conf.level = 0.95,
         method = "spearman")
```


```{r}
#Correlations: solution index and CA1, CS23, MTL, totalhipp

########## Spearman for solution index

#CA1
cor.test(data_women$t2hipp_vol_ca1_cor,
         data_women$dsp_si_true_pct,
         conf.level = 0.95,
         method = "spearman")

#CA23
cor.test(data_women$t2hipp_vol_ca23_cor,
         data_women$dsp_si_true_pct,
         conf.level = 0.95,
         method = "spearman")

#MTL
cor.test(data_women$t2hipp_vol_mtl_cor,
         data_women$dsp_si_true_pct,
         conf.level = 0.95,
         method = "spearman")

#Total hipp
cor.test(data_women$t2hipp_vol_totalhipp_cor,
         data_women$dsp_si_true_pct,
         conf.level = 0.95,
         method = "spearman")
```






























