---
title: "snag_wrangling_tests"
author: "Shuying Yu"
date: "3/11/2022"
output: html_document
---


```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center")

#Attach libraries
library(tidyverse)
library(here)
library(janitor)
library(ggpubr)
library(patchwork)
```


# Power Analyses

https://davidakenny.shinyapps.io/MedPower/

## Study 2

X = FSH, Y = DSP solution index, M = CA1 volume

Data has been cleaned into tidy format prior to analysis.

```{r data_wrangling}
########## Data 

#Read in the data
df <- read_csv(here("data", "snag_t1_t2hipp_hormones.csv"))



#Filter for having T1 and scan hormone data
df_t1 <- df %>% 
  
  #Filter
  filter(t1_status == "yes") %>% 
  
  #Select certain columns
  select(subject_id, sex, age_scan_years,
         repo_status_scan, t1_vol_left_hipp:t1_csf_vbm,
         loop_pe_rad1_m:dsp_si_true_pct,
         estradiol_scan_pg_ml:testosterone_scan_ng_dl)


  
df_t2hipp <- df %>% 
  
  #Filter
  filter(t2hipp_status == "yes") %>% 
  
  #Select certain columns
  select(subject_id, sex, age_scan_years,
         repo_status_scan, t1_tiv_vbm,
         t2hipp_tiv_thick_cat12:t2hipp_thick_right_phg,
         loop_pe_rad1_m:dsp_si_true_pct,
         estradiol_scan_pg_ml:testosterone_scan_ng_dl)




########## Correct using TIV, then average left and right

##### T1
df_t1_corrected <- df_t1 %>% 

  #Correct using TIV from VBM
  #(Total) hippocampus, MTL, parahippocampal gyrus
  #Include total GMV from hippocampus 
  mutate(t1_vol_left_hipp = t1_vol_left_hipp/t1_tiv_vbm,
         t1_vol_right_hipp = t1_vol_left_hipp/t1_tiv_vbm,
         t1_vol_left_mtl = t1_vol_left_mtl/t1_tiv_vbm,
         t1_vol_right_mtl = t1_vol_right_mtl/t1_tiv_vbm,
         t1_vol_left_parahipp = t1_vol_left_parahipp/t1_tiv_vbm,
         t1_vol_right_parahipp = t1_vol_right_parahipp/t1_tiv_vbm,
         t1_vol_gmv = t1_gmv_vbm/t1_tiv_vbm) %>% 
  
  #Average left and right regions
  mutate(t1_vol_hipp = (t1_vol_left_hipp + t1_vol_right_hipp)/2,
         t1_vol_mtl = (t1_vol_left_mtl + t1_vol_right_mtl)/2,
         t1_vol_parahipp = (t1_vol_left_parahipp + t1_vol_right_parahipp)/2)



##### T2hipp

#Create function for apply to variables
dividebyTIV <- function(x, na.rm = FALSE) (x/df_t2hipp$t1_tiv_vbm)

#Apply function to variables
df_t2hipp_corrected <- df_t2hipp %>% 
  
  #Correct using TIV from T1 VBM
  #CA1, CA23, DG, ERC,PHC, PRC, SUB, PHG
  mutate_at(vars(t2hipp_vol_left_ca1:t2hipp_vol_right_phg),
            dividebyTIV) %>% 
  
  #Average left and right regions
  mutate(t2hipp_vol_ca1 = (t2hipp_vol_left_ca1 + t2hipp_vol_right_ca1)/2,
         t2hipp_vol_ca23 = (t2hipp_vol_left_ca23 + t2hipp_vol_right_ca23)/2,
         t2hipp_vol_dg = (t2hipp_vol_left_dg + t2hipp_vol_right_dg)/2,
         t2hipp_vol_erc = (t2hipp_vol_left_erc + t2hipp_vol_right_erc)/2,
         t2hipp_vol_phc = (t2hipp_vol_left_phc + t2hipp_vol_right_phc)/2,
         t2hipp_vol_sub = (t2hipp_vol_left_sub + t2hipp_vol_right_sub)/2,
         t2hipp_vol_phg = (t2hipp_vol_left_phg + t2hipp_vol_right_phg)/2)
```


### Correlation FSH and CA1 volume

```{r}
cor.test(df_t2hipp_corrected$fsh_scan_miu_ml, 
         df_t2hipp_corrected$t2hipp_vol_ca1,
         method = "pearson")
```

### Correlation FSH and DSP Solution Index

```{r}
cor.test(df_t2hipp_corrected$fsh_scan_miu_ml, 
         df_t2hipp_corrected$dsp_si_true_pct,
         method = "pearson")
```

### Correlation CA1 and DSP Solution Index

```{r}
cor.test(df_t2hipp_corrected$dsp_si_true_pct, 
         df_t2hipp_corrected$t2hipp_vol_ca1,
         method = "pearson")
```

Effect	Beta	Partial r	Power	N
c (total)	-.698	-.698	virtually 1	61
a	-.370	-.370	.853	61
b	-.250	-.324	.728	61
c' (direct)	-.790	-.735	virtually 1	61
ab (indirect)	.092		.621	61
Alpha for all power calculations set to .050. Effects (a, b, and c') are Betas.




